Quantifying Spatially Resolved Hydration Thermodynamics Using Grid Inhomogeneous Solvation Theory [Article v\versionnumber]

### 2022-12-07 manuscript/manuscript.tex
Move MD part to basic tutorial, other minor edits

\subsubsection{Equilibration}
Equilibrate your solvated biotin in an NVT or NpT ensemble.
Even when running the simulation in NVT, it is advisable to include an NpT equilibration step to adjust the box size beforehand.
Use enough equilibration time to allow the water to relax around your biotin structure.

\subsubsection{MD simulation}
\todo{Move simulation part from complex to here}

@@

Use the \software{equilibration.py} script (or your own) to perform short NVT and NpT equilibration runs on the complex structure.
Note that the equilibration script contains unrestrained NpT steps, because restraining two molecules at the same time might interfere with the pressure equilibration.

Then split the equilibrated structure into biotin and streptavidin parts using \software{cpptraj\_remove\_mol.sh}.
After that, you can use \software{equilibration.py -R} to run short restrained 
equilibrations (where only the water is allowed to move) on the individual 
systems.

\subsubsection{Running MD for GIST}

Now, you can run restrained MD simulations based on the equilibrated structures for subsequent GIST analyses.
Restraints should be used during the MD because IST theory is not defined for moving solutes.
We apply a harmonic restraint of \SI{100}{\kilo\calorie\per\mole\per\angstrom\squared} to all heavy atoms to keep the system in place.
We run \SI{100}{\nano\second} of simulation and store the conformations every \SI{100}{\pico\second}.
A simulation length of at least \SI{20}{\nano\second} is recommended to obtain statistically independent snapshots of the water molecules.

\subsection{GIST analysis of Streptavidin-Biotin}
For easier post-processing, here we use grids that contain the whole molecule.
Smaller grids encompassing only the binding site could be used for streptavidin and the complex, but this would complicate the analysis.
%This could in theory be avoided for streptavidin and the complex, since only the surrounding of the biotin molecule (i.e., the binding pocket) is required for the analysis.
Furthermore, we recommend that you center the molecule in space.
For each system (biotin, streptavidin, complex), produce a centered pdb file using \software{cpptraj}:
\begin{lstlisting}[style=cpptraj]
parm solvated.parm7
trajin md-01.nc 1 1 1
autoimage !(:WAT) origin
strip :WAT
trajout solute-centered.pdb
\end{lstlisting}

@@

\begin{lstlisting}[style=python]
difference = compl_rdf - biotin_rdf - strept_rdf
cutoff = 12
integral = difference[bins < cutoff].sum()
print("Integral = {}".format(integral))
plt.plot(bins, np.cumsum(difference))
plt.axvline(cutoff)
plt.xlabel('distance to biotin [Å]')
plt.ylabel('dG contribution [kcal/mol]')
\end{lstlisting}
% \todo{\$\textbackslash AA\$ should produce an angstrom sign in matplotlib (what a pain to have this printed out in \LaTeX\ tho) -vah} -- Å produces an angstrom sign ;-)
Compute the energy (\inlinecode{Eall}) and entropy (\inlinecode{dTSsix}) contributions separately.
You will find that the energy is strongly unfavorable to binding, because we do not yet include the interaction energy between biotin and streptavidin.

Using the \inlinecode{energy} command in \software{cpptraj}, you can compute this energy difference.
We recommend using PME in combination with PME-GIST, but not with GPU-GIST\@.
An example \software{cpptraj} input might look like:
\begin{lstlisting}[style=cpptraj]
parm solvated.parm7
trajin md-01.nc 1 last 100
energy complex ^1,2 # etype pme
energy strept ^1 # etype pme
energy biotin ^2 # etype pme
go
diff = complex[total] - strept[total] - biotin[total]
writedata energy.dat diff complex[total] \
strept[total] biotin[total]
avg(complex[total])
avg(strept[total])
avg(biotin[total])
avg(diff)
\end{lstlisting}

It has been shown in literature \cite{Chen2021,Waibl2022-gist-solvents} that the total solvation entropy in water is best approximated by 0.6 times the first order entropy.
Also compute the scaled entropy and check the effect on the binding affinity.
The expected results are summarized in Table~\ref{tab_dg_monomers_dimer}.
%\todo{Table 1 is placed quite akwardly at the moment for me, maybe enforce its position? -vah. Ok? -fwa}
\todo{this table is super confusing to me. What's "total"? Doesn't seem to be E-TS. Diff is also not the sum of other total entries. Shouldn't it be TS instead of S?-vah}
\begin{table}[h]
	\caption{Free energy contributions for the monomers and the dimer in kcal$\cdot$mol$^{-1}$. ``total'' is calculated as $E^\text{internal} + \Delta E^\text{GIST} - \Delta S^\text{scaled}$.}\label{tab_dg_monomers_dimer}
	\small
	\begin{tabular}{@{}lrrrrr@{}}
		\toprule
		System       & E(internal) & $\Delta E^\textit{GIST}$ & $\Delta S^\textit{GIST}$ & $\Delta S^\textit{scaled}$ & total \\
		\midrule
		complex      & -1303.4 & -361.2 & -233.1 & -139.9 & -1524.7 \\
		streptavidin & -1180.7 & -360.8 & -242.4 & -145.4 & -1396.1 \\
		biotin       & -24.0   &  -98.1 &  -35.1 &  -21.1 &  -101.0 \\
		Diff         & -98.7   &   97.7 &   44.4 &   26.6 &   -27.6 \\
		\bottomrule
	\end{tabular}
\end{table}

>>>>

\subsubsection{Equilibration}
Equilibrate your solvated biotin in an NVT or NpT ensemble.
Even when running the simulation in NVT, it is advisable to include an NpT equilibration step to adjust the box size beforehand.
Use enough equilibration time to allow the solvent to relax. 
For water, \SI{0.5}{\nano\second} are usually sufficient to equilibrate the density, but narrow binding pockets can take longer.

\subsubsection{MD simulation}
Now, you can run restrained MD simulations based on the equilibrated structures.
Restraints should be used because IST theory is not defined for moving solutes.
We apply a harmonic restraint of \SI{100}{\kilo\calorie\per\mole\per\angstrom\squared} to all heavy atoms to keep the system in place.
We run \SI{100}{\nano\second} of simulation and store the conformations every \SI{100}{\pico\second}.
A simulation length of at least \SI{20}{\nano\second} is recommended to obtain statistically independent snapshots of the water molecules.

@@

Use the \software{equilibration.py} script (or your own) to perform short NVT and NpT equilibration runs on the complex structure.
It is important to include unrestrained NpT steps, because restraining two molecules at the same time might interfere with the pressure equilibration.

Then split the equilibrated structure into biotin and streptavidin parts using \software{cpptraj\_remove\_mol.sh}.
After that, use \software{equilibration.py -R} to run short restrained 
equilibrations (where only the water is allowed to move) on the individual 
systems.
By using restraints, we can make sure that the monomer conformations match the complex.

\subsubsection{MD Simulation}
The MD can be run as described above for each monomer and the complex.

\subsubsection{GIST analysis of Streptavidin-Biotin}
For easier post-processing, here we use grids that contain the whole molecule.
Smaller grids encompassing only the binding site could be used for streptavidin and the complex, but this would complicate the analysis.
%This could in theory be avoided for streptavidin and the complex, since only the surrounding of the biotin molecule (i.e., the binding pocket) is required for the analysis.
Furthermore, we recommend that you center the molecule in space.
For each system (biotin, streptavidin, complex), produce a centered pdb file using \software{cpptraj}:
\begin{lstlisting}[style=cpptraj]
parm solvated.parm7
trajin md-01.nc 1 1 1
autoimage !(:WAT) origin
strip :WAT
trajout solute-centered.pdb
\end{lstlisting}

@@

\begin{lstlisting}[style=python]
difference = compl_rdf - biotin_rdf - strept_rdf
cutoff = 12
integral = difference[bins < cutoff].sum()
print("Integral = {}".format(integral))
plt.plot(bins, np.cumsum(difference))
plt.axvline(cutoff)
plt.xlabel('distance to biotin [Å]')
plt.ylabel('dG contribution [kcal/mol]')
\end{lstlisting}
% \todo{\$\textbackslash AA\$ should produce an angstrom sign in matplotlib (what a pain to have this printed out in \LaTeX\ tho) -vah} -- Å produces an angstrom sign ;-)
Compute the energy (\inlinecode{Eall}) and entropy (\inlinecode{dTSsix}) contributions separately.
You will find that the energy disfavors binding, because we do not yet include the interaction energy between biotin and streptavidin.

Using the \inlinecode{energy} command in \software{cpptraj}, you can compute this interaction energy.
We recommend using PME in combination with PME-GIST, but not with GPU-GIST\@.
An example \software{cpptraj} input might look like:
\begin{lstlisting}[style=cpptraj]
parm solvated.parm7
trajin md-01.nc 1 last 100
energy complex ^1,2 # etype pme
energy strept ^1 # etype pme
energy biotin ^2 # etype pme
go
diff = complex[total] - strept[total] - biotin[total]
writedata energy.dat diff complex[total] \
strept[total] biotin[total]
avg(complex[total])
avg(strept[total])
avg(biotin[total])
avg(diff)
\end{lstlisting}

It has been shown \cite{Chen2021,Waibl2022-gist-solvents} that the solvation entropy in water is best approximated by 0.6 times the first order entropy.
Also compute the scaled entropy and check the effect on the binding affinity.
The expected results are summarized in Table~\ref{tab_dg_monomers_dimer}.
%\todo{Table 1 is placed quite akwardly at the moment for me, maybe enforce its position? -vah. Ok? -fwa}
\todo{this table is super confusing to me. What's "total"? Doesn't seem to be E-TS. Diff is also not the sum of other total entries. Shouldn't it be TS instead of S?-vah}
\begin{table}[h]
	\caption{Free energy contributions for the monomers and the dimer in kcal$\cdot$mol$^{-1}$. ``Diff'' is calculated as $E^\text{internal} + \Delta E^\text{GIST} - \Delta S^\text{scaled}$.}\label{tab_dg_monomers_dimer}
	\small
	\begin{tabular}{@{}lrrrrr@{}}
		\toprule
		System       & E(internal) & $\Delta E^\textit{GIST}$ & $T\Delta S^\textit{GIST}$ & $T\Delta S^\textit{scaled}$ & Diff \\
		\midrule
		complex      & -1303.4 & -361.2 & -233.1 & -139.9 & -1524.7 \\
		streptavidin & -1180.7 & -360.8 & -242.4 & -145.4 & -1396.1 \\
		biotin       & -24.0   &  -98.1 &  -35.1 &  -21.1 &  -101.0 \\
		Diff         & -98.7   &   97.7 &   44.4 &   26.6 &   -27.6 \\
		\bottomrule
	\end{tabular}
\end{table}
